An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-08-05 EconGrowth2.Rnw
Edits to the paper.

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
for each economy (see Section~\ref{sec:Economy_Types}) and 
each energy quantification (see Section~\ref{sec:Quantification_of_Energy}) 
via least-squares methods. We used \texttt{R} \citep{R} functions
\texttt{nls} and \texttt{cesEst} \citep{Henningsen:2011td} 
for that purpose. The following subsections discuss 
parameter estimation methods and results.

@@

When resampling, the modeler assumes that the residuals ($r_t$) between 
observed data (in our case historical GDP values, $y_{act,t}$) and
model predictions (in our case, $y_{pred,t}$) at
time $t$ are representative 
of the uncertainty and random scatter in the observed historical data.

\begin{equation} \label{eq:residual}
  r_t = y_{pred,t} - y_{act,t}
\end{equation}

Under that assumption, it would not be surprising if the residual from 
any one year had been observed on a different year. 
By applying randomly selected residuals (with randomly selected signs)
to model predictions at each year of the GDP time series, we create a resampled
GDP time series ($y_{resample,t}$) that can be re-fitted using the techniques
described below.

@@

The coefficients from the fit to the resampled time series
(the ``resample coefficients'') will be different from the coefficients 
obtained from the fit to historical data (the ``base coefficients'').
Large residuals in the base fit
are expected to produce large variation in resample coefficients, 
indicating less certainty in the knowledge
of those coefficients. 
Similarly, small base fit residuals are expected to produce small
variation in resample coefficients, indicating better knowledge
of the model coefficients.

It is important to note that large residuals can arise from 
(a) poor quality historical time series data ($y_{act}$) or 
(b) a mathematical model that does not describe the underlying phenomena well.
In either case, we would not want to claim precise knowledge of the 
fitted coefficients for a model. The resampling technique that we employed
provides looser coefficient distributions when either of the above
is true. After discussing the fitting technique for each 
growth model, we show scatter plots of fitted coefficients
to provide visual indication of the uncertainty 
of the coefficients for each model.

>>>>

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
for each energy quantification (see Section~\ref{sec:Quantification_of_Energy}) 
and each economy (see Section~\ref{sec:Economy_Types})
via least-squares methods. We used \texttt{R} \citep{R} functions
\texttt{nls} and \texttt{cesEst} \citep{Henningsen:2011td} 
for that purpose. The following subsections discuss 
parameter estimation methods and results.

@@

When employing resampling techniques, 
the modeler assumes that the residuals ($r_t$) between 
observed data (in our case historical GDP values, $y_{act,t}$) and
model predictions (in our case, $y_{pred,t}$) at
time $t$ are representative 
of the uncertainty of and random scatter in the observed historical data.

\begin{equation} \label{eq:residual}
  r_t = y_{act,t} - y_{pred,t}
\end{equation}

Under that assumption, it would not be surprising if the residual from 
any one year had been observed on a different year. 
By applying randomly selected residuals (with randomly selected signs)
to model predictions at each year of the GDP time series, we create a resampled
GDP time series ($y_{resample,t}$) that can be fitted using the techniques
described below.

@@

The coefficients from the fit to the resampled time series
(the ``resample coefficients'') will be different from the coefficients 
obtained from the fit to historical data (the ``base coefficients'').
Large residuals in the fit to historical data (the ``base fit'')
are expected to produce large variation in resample coefficient values
(when the resampling is repeated \Sexpr{numResamples()} times),
indicating less certainty in the value
of those coefficients. 
Similarly, small base fit residuals are expected to produce small
variation in resample coefficient values, indicating more-precise knowledge
of the model coefficients.

It is important to note that large residuals can arise from 
(a) poor quality historical time series data ($y_{act}$) or 
(b) a mathematical model that does not describe the underlying phenomena well.
In either case, we would not want to claim precise knowledge of the 
fitted coefficients for an economic growth model. 
The resampling technique that we employed
provides looser coefficient distributions when either of the above
is true. After discussing the fitting technique for each 
growth model below, we show scatter plots of fitted coefficients,
thereby providing visual indication of the uncertainty 
of the coefficients for each model.
